NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-01-31 SI.tex
Update authors
Update affiliations

\author[1]{Anne Kiirikki}         %ORCID: ?
\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
\author[2]{Lara Bort}             %ORCID: ?
\author[3]{Pavel Buslaev}         %ORCID: ?
\author[4]{Fernando Favela}
\author[5]{Tiago Mendes Ferreira}
\author[6,7]{Patrick Fuchs}
\author[8,9]{Rebeca Garcia-Fandino}
\author[2]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[2]{Batuhan Kav}           %ORCID: ?
\author[2]{Milla Kurki}
\author[2]{Alexander Kuzmin}
\author[10,11]{Jesper Madsen}
\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[12]{Thomas Piggot}
\author[2]{Suman Samantray}
\author[2]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}

@@

\begin{abstract}
%We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
\end{abstract}

>>>>

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[3]{Lara Bort}             %ORCID: ?
\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[5]{Fernando Favela}
\author[6]{Tiago Mendes Ferreira}
\author[7,8]{Patrick F.J. Fuchs}
\author[9]{Rebeca Garcia-Fandino}
\author[10]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[13]{Patrik Kula}
\author[14]{Milla Kurki}
\author[10]{Alexander Kuzmin}
\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[19]{Thomas Piggot}
\author[20]{{\'A}ngel Pi{\~n}eiro}
\author[10]{Suman Samantray}
\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
%\affil[2]{Affiliation, department, city, postcode, country}
\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[10]{no affiliation}
\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[17]{Department of Chemistry, University of Bergen, Norway}
\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}

@@

%\begin{abstract}
%We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
%\end{abstract}

### 2023-01-31 main.tex
Update affiliations

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[14]{Lara Bort}             %ORCID: ?
\author[3]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[4]{Fernando Favela}
\author[5]{Tiago Mendes Ferreira}
\author[6,7]{Patrick F.J. Fuchs}
\author[8]{Rebeca Garcia-Fandino}
\author[2]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[2]{Batuhan Kav}           %ORCID: ?
\author[13]{Patrik Kula}
\author[2]{Milla Kurki}
\author[2]{Alexander Kuzmin}
\author[10,11]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[12]{Thomas Piggot}
\author[9]{{\'A}ngel Pi{\~n}eiro}
\author[2]{Suman Samantray}
\author[8,9,15]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[7]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[9]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[10]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[11]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[13]{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
\affil[15]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}

@@

\begin{abstract}
Cellular membrane lipid composition is implicated in diseases and major biological functions in cells, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour. Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer inaccuracies in model parameters. Applications of data-driven and machine learning methods, currently revolutionising many fields, are limited to membrane systems due to the lack of suitable training sets. Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers. The NMRlipids databank will benefit scientists in different disciplines by provoding automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface. To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.

@@

\subsection{Applications of the NMRlipids databank}
\subsubsection{Selection of simulation parameters using the NMRlipids databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models. However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}. 
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
For example, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}. We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations. To streamline this process,
%the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters,
we have defined quantitative quality measures, based on C--H bond order parameters from NMR and X-ray scattering form factors, for lipid bilayer simulations in the NMRlipids databank. 
%The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments. 
The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}. For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
%For detailed definitions of quality measures, see the supplementary information.
Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.

@@

\subsection{Principal Components Analysis of equilibration of simulations}
To estimate how well the provided trajectories are converged, the Principal Components Analysis (PCA) was used \cite{buslaev16,buslaev2020principal}. Following the PCALipids \cite{buslaev2020principal} protocol, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. Next, each lipid configuration was aligned to the average lipid structure. Afterwards, the PCA was applied on the cartesian coordinates of all heavy atoms of the lipid. Since it was shown before, that the motions along the first, major, principal component (PC) are the slowest ones \cite{buslaev16}, the equilibration of individual lipids was estimated based on the dynamics of the first PC. The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}. However, due to the linear dependence between autocorrelation decay times and distribution convergence times, faster calculation times and higher computational stability of autocorrelation decay times\cite{buslaev16, buslaev2020principal}, the distribution convergence timescales were calculated from autocorrelation decay times:
\begin{equation}
    \tau_{\text{convergence}} = 49*\tau_{\text{autocorrelation}}
\end{equation}

The empirical coefficient 49 was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields. All simulations, used for calculating the empirical coefficient, that relate convergence and autocorrelation, were at least 200 ns long. As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations as the most frequent in the Databank were used.

@@

P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
R.G.-F. thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09). 
AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00). 
O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.
\section*{Author contributions statement}
%Must include all authors, identified by initials, for example:
%A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

>>>>

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[3]{Lara Bort}             %ORCID: ?
\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[5]{Fernando Favela}
\author[6]{Tiago Mendes Ferreira}
\author[7,8]{Patrick F.J. Fuchs}
\author[9]{Rebeca Garcia-Fandino}
\author[10]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[13]{Patrik Kula}
\author[14]{Milla Kurki}
\author[10]{Alexander Kuzmin}
\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[19]{Thomas Piggot}
\author[20]{{\'A}ngel Pi{\~n}eiro}
\author[10]{Suman Samantray}
\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
%\affil[2]{Affiliation, department, city, postcode, country}
\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[10]{no affiliation}
\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[17]{Department of Chemistry, University of Bergen, Norway}
\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}

@@

\begin{abstract}
Cellular membrane lipid composition is implicated in diseases and major biological functions in cells, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour. Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer inaccuracies in model parameters. Applications of data-driven and machine learning methods, currently revolutionising many fields, are limited to membrane systems due to the lack of suitable training sets. Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers. The NMRlipids databank will benefit scientists in different disciplines by providing automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface. To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.

@@

\subsection{Applications of the NMRlipids databank}
\subsubsection{Selection of simulation parameters using the NMRlipids databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models. However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}. 
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
For example, Fig.~\ref{fig:POPCPOPEapls}C illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}. We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations. To streamline this process,
%the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters,
we have defined quantitative quality measures, based on C--H bond order parameters from NMR and X-ray scattering form factors, for lipid bilayer simulations in the NMRlipids databank. 
%The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments. 
The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}. For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
%For detailed definitions of quality measures, see the supplementary information.
Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.

@@

\subsection{Principal Components Analysis of equilibration of simulations}
To estimate how well the provided trajectories are converged, the Principal Components Analysis (PCA) was used \cite{buslaev16,buslaev2020principal}. Following the PCALipids \cite{buslaev2020principal} protocol, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. Next, each lipid configuration was aligned to the average lipid structure. Afterwards, the PCA was applied on the cartesian coordinates of all heavy atoms of the lipid. Since it was shown before, that the motions along the first, major, principal component (PC) are the slowest ones \cite{buslaev16}, the equilibration of individual lipids was estimated based on the dynamics of the first PC. The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}. However, due to the linear dependence between autocorrelation decay times and distribution convergence times, faster calculation times and higher computational stability of autocorrelation decay times\cite{buslaev16, buslaev2020principal}, the distribution convergence timescales were calculated from autocorrelation decay times:
\begin{equation}
    \tau_{\text{convergence}} = k*\tau_{\text{autocorrelation}}
\end{equation}

The empirical coefficient $k$ is equal to $k = 49$ and was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields. All simulations, used for $k$ calculation, were at least 200 ns long. As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations, as the most frequent in the Databank, were used.

@@

P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
R.G.-F. thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09). 
AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00). 
M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.
\section*{Author contributions statement}
%Must include all authors, identified by initials, for example:
%A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
